---
title: "ESM 244 Lab 1"
author: "Indigo Bannister"
date: "1/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(sf)
library(tmap)

```

### Read in sf trees data
```{r}
sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"))

```

### Basic wrangling reminders

Find the top 5 highest observations of trees by legal_status, do some wrangling.

```{r}
top_5_status <- sf_trees %>% 
  count(legal_status) %>% 
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
  relocate(tree_count) %>% 
  slice_max(order_by = tree_count, n = 5)
  
```

Make a graph of the top 5 trees by legal status.
```{r}
ggplot(data = top_5_status, aes(x = fct_reorder(legal_status, tree_count), y = tree_count)) +
  geom_col() +
  labs(x = "Legal Status", y = "Tree Count") +
  coord_flip() +
  theme_minimal()

```

### A few more data wrangling refresher examples

Only want to keep observations for blackwood acacia trees.

```{r}
blackwood_acacia <- sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>% 
  select(legal_status, date, latitude, longitude)

```

Make a fake map (r will not know that lat/long are coordinates, just using x/y)
```{r}
ggplot(data = blackwood_acacia, aes(x = longitude, y = latitude)) + 
  geom_point()
```

### Tidyr::seperate() and unite() functions
Useful for combining or separating columns

```{r}
sf_trees_sep <- sf_trees %>% 
  separate(species, into = c("spp_scientific", "spp_common"), sep = "::")

sf_trees_unite <- sf_trees %>% 
  unite("id_status", tree_id:legal_status, sep = "_cool!_")
```

### Make some actual maps of blackwood acacia trees in SF

`st_as_sf()` to convert lat and long to spatial coordinates

```{r}
blackwood_acacia_sp <- blackwood_acacia %>% 
  drop_na(longitude,latitude) %>% 
  st_as_sf(coords = c("longitude", "latitude"))

st_crs(blackwood_acacia_sp) = 4326

ggplot(data = blackwood_acacia_sp) +
  geom_sf(color = "darkgreen")
```

Read in SF roads shapefile:
```{r}
sf_map <- read_sf(here("data","sf_map", "tl_2017_06075_roads.shp"))

# run st_transform in console to see what projection is

st_transform(sf_map, 4326)

ggplot(data = sf_map) +
  geom_sf()


```

Combine the blackwood acacia and SF roads map

```{r}
ggplot() +
  geom_sf(data = sf_map, size = 0.1, color = "darkgray") +
  geom_sf(data = blackwood_acacia_sp, color = "darkgreen") + 
  theme_minimal()
```

Now to create an interactive map:
```{r}
tmap_mode("view")

tm_shape(blackwood_acacia_sp) + 
  tm_dots()
```

